### 2024-02-15 other/ssrp/ssrp_report.qmd
Wrote ssrp

---
title: "SSRP Report based on *Energy Saving May Kill: Evidence from the Fukushima Nuclear Accident* by He and Tanaka (2019)"
author:
  - Yi Fei Pang
  - Irene Huynh
  - Tara Chakkithara
thanks: "Code and data are available at: https://github.com/apang00/Energy-Policy.git"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

# Report

The statistical programming language R was used to acquire, clean, and process the data [@citeR]. In particular, the following R packages were used: tidyverse [@tidyverse], readxl [@readxl], haven [@haven], ggthemes [@ggthemes], and ggplot2 [@ggplot2].

Three figures from *Energy Saving May Kill: Evidence from the Fukushima Nuclear Accident* by He and Tanaka (2019) were reproduced:

- Figure 1 is a reproduction of Figure 1A. from He and Tanaka (2019).
- Figure 2 is a reproduction of Figure 1B. from He and Tanaka (2019).
- Figure 3 is a reproduction of Figure G1. from He and Tanaka (2019).

![](../../other/figures/figure1.png)

![](../../other/figures/figure2.png)
![](../../other/figures/figure3.png)
\newpage

# References